risk-perception

### 2019-10-23 risk-percption.Rmd
lpsolve code

```{r}
wheat_dry <- 0
wheat_wet <- 1.6
barley_dry <- 0.93
barley_wet <- 1.18

calc_payoff <- function(wheat_prop, risk) {
  barley <- (1 - wheat_prop) * (barley_dry * risk + barley_wet * (1 - risk))
  wheat <- wheat_prop * (wheat_dry * risk + wheat_wet * (1 - risk))
  barley + wheat
}

tibble(wheat_prop)
```

calc_ratio <- function(risk) {
  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
}
calc_ratio(.2)
```

So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.
```{r}
test_dat <- expand_grid(wheat_prop = (0:10)/10,
       risk = seq(0, 1, 0.2)) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk)+ barley_dry * risk),
         payoff = calc_payoff(wheat_prop, risk))

@@

test_dat %>%
  select(-(util_dry:util_wet)) %>%
  gather(type, util, util_wheat:util_barley) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)
test_dat %>%
  select(-(util_wheat:util_barley)) %>%
  gather(type, util, util_dry:util_wet) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)

@@

@@

>>>>

```{r}
wheat_dry <- 0
wheat_wet <- 1.6
barley_dry <- 0.93
barley_wet <- 1.18
```

```{r}
library(lpSolve) #Loading lpSolve library
risk <- 0.6
wheat_exp <- wheat_wet * (1 - risk) + wheat_dry * risk
barley_exp <- barley_wet * (1 - risk) + barley_dry * risk
obj.fun <- c(wheat_exp, barley_exp)

constr=matrix(c(1,1,1,0,0,1,wheat_exp,barley_exp),ncol = 2, byrow = TRUE) #Loading the constraints
constr.dir=c("=",">=",">=",">=")
constr.rhs=c(1,0,0,1)

mod=lp("max",obj.fun,constr,constr.dir,constr.rhs,compute.sens = TRUE) #Using lp() to solve our problem

mod$solution
```

```{r}
calc_payoff <- function(x, drought) {
  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet

  x %>%
    mutate(yield = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop))
}

update_prop <- function(x, aspiration) {
  x %>%
    mutate(new_prop = round(runif(n()), 2),
           wheat_prop = if_else(yield > aspiration, wheat_prop, new_prop)) %>%
    select(-new_prop)
}

test3 <- tibble(agent = 1:1000) %>%
  mutate(wheat_prop = round(runif(n()), 2)) %>%
  calc_payoff(drought = TRUE)
test3 %>% update_prop(aspiration = .6)

accumulate((runif(25) < .3), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .5), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(as.numeric(time), yield, group = agent)) + geom_line(alpha = .2)

```{r}
library(ggridges)
accumulate((runif(25) < .2), ~calc_payoff(.x, drought = .y) %>% update_prop(aspiration = .6), .init = test3) %>% bind_rows(.id = 'time') %>% ggplot(aes(wheat_prop, y = as.numeric(time), group = as.numeric(time))) +
  geom_density_ridges() +
  theme_minimal()
```

```{r}
calc_ratio <- function(risk) {
  (risk * barley_dry + (risk - 1) * barley_wet) / (risk * (barley_dry - wheat_dry) + wheat_wet * -1 * risk + wheat_wet + barley_wet * (risk - 1))
}
calc_ratio(.2)
```

So at between 3 in 10 and 4 in 10 risk of drought, a maximizer will switchfrom wheat to barley.

```{r}
tibble(wheat_prop = (0:10)/10,
       risk = 0.5,
       util_wet = wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop),
      util_dry = wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop),
      util_expected = util_wet * (1 - risk) + util_dry * risk,
      min_yield = pmin(util_wet, util_dry)) %>%
    ggplot(aes(util_expected, min_yield)) + geom_line() +
  geom_hline(yintercept = .6, linetype = 2)
```

```{r}
tibble(wheat_prop = (0:10)/10,
       risk = 0.5,
       util_wet = wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop),
      util_dry = wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop),
      util_expected = util_wet * (1 - risk) + util_dry * risk) %>%
    gather(type, util, util_wet:util_expected) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line()
```

```{r}
test_dat <- expand_grid(wheat_prop = (0:10)/10,
       risk = seq(0, 1, 0.2)) %>%
  mutate(util_wet = (1 - risk) * (wheat_wet * wheat_prop + barley_wet * (1 - wheat_prop)),
         util_dry = risk * (wheat_dry * wheat_prop + barley_dry * (1 - wheat_prop)),
         util_wheat = wheat_prop * (wheat_wet * (1 - risk) + wheat_dry *  risk),
         util_barley = (1 - wheat_prop) * (barley_wet* (1 - risk)+ barley_dry * risk),
         payoff = calc_payoff(wheat_prop, risk))

test_dat %>%
  select(-(util_wheat:util_barley)) %>%
  gather(type, util, c(util_dry, util_wet, payoff)) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)

@@

test_dat %>%
  select(-(util_dry:util_wet)) %>%
  gather(type, util, util_wheat:util_barley) %>%
    ggplot(aes(wheat_prop, util, color = type)) + geom_line() + facet_wrap(~risk)

@@

```{r}
tibble(x)
```

@@

plot(seq(.1,1,.1) * (1 - seq(.1,1,.1)))
